Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-05-29 WRITEUP/ms.tex
updated ref

Summarizing, the virial parameter can be used to quantify the stability/boundedness of
%mm an MCC.
    a molecular cloud.
Accounting for the external pressure, a virial parameter of $\alpha_{\rm vir}\lesssim2$ would be unstable \citep{bertoldi:1992}.
%
Such a criterion is often used in \obs \citep[see e.g., ][]{Kauffmann17b}.

@@

In our stability analysis, we account for the combined effect of gas and stars \citep[derived exactly by][]{Rafikov01}, and for the non-negligible disk thickness. This is done by adopting an approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modifies the $Q$ parameter for gas and stars by accounting for the vertical velocity dispersion
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the finite disk thickness reduces the gravity in the vertical direction, thereby making it easier for a system to maintain stability, and thus lowering the critical Toomre $Q_{\rm crit}$ from $\simeq$\,1 to 0.67 \citep{Goldreich65a}.
%
On the other hand, including the contribution of the stellar component promotes gravitational instability, and thus increases
$Q_{\rm crit}$, more-so if the stars have low velocity dispersion.
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

@@

Star formation is modeled using an H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar tracks from \ncode{starburst99} and include stellar feedback from supernovae (SNe) as well as winds from OB and asymptotic giant branch stars. Coupling to the gas is implemented via a sub-grid model for blastwaves, which accounts for radiative energy losses inside the cell. The remaining energy is injected into the ISM in kinetic and thermal energy forms. Radiation pressure on dust and gas is also included. Photoionization from local sources is neglected in this simulation \citep[see][for its effect]{pallottini:2019,decataldo:2019}.
%
Via feedback processes the gas can acquire energy both thermal and non-thermal form \citep{agertz:2013}; the latter is a term that mimics unresolved turbulence: such non-thermal energy is not affected by cooling but rather is dissipated on the eddy turn-over time scale \citep{maclow1999turb}. See \citet{Pallottini17b} for further details.

The simulation zooms in on a galaxy named after the flower \flower whose properties are given in \citet{Pallottini17a} and are briefly summarized in the following. \flower is a Lyman-break galaxy that at $z\simeq 6$ is hosted by a dark matter halo of mass $M_{\rm DM}\simeq$\,10$^{11}$\,\Msun at the center of a cosmic web knot, and accretes mass from the IGM mainly via three filaments of length $\simeq$\,100\,kpc. \flower has a stellar mass of $M_\star\simeq$\,3\E{10}\,\Msun, a metallicity of $Z\simeq$\,0.5\,$Z_{\odot}$, a molecular gas mass of $M_{\rm H2}\simeq$\,5\E{7}\,\Msun, and a SFR of 30\,--\,80\,\Msun\,yr\pmOne.
%
The specific SFR of \flower is ${\rm sSFR} \equiv {\rm SFR}/M_{\star} \simeq$~4--40 ~Gyr\pmOne, compatible with the sSFR vs.\ $M_{\star}$ relation observed at high redshift \citep{Jiang2016}, which is higher than normal local galaxies. The effective stellar radius of \flower is $\sim$0.5\,kpc and the dark matter virial radius is $r_{\rm 200}\simeq$\,15\,kpc.

@@

The high pressure observed in \flower (see \citealt{Pallottini17a}) is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{kohandel:2019} and higher velocity accretion/SN-driven outflows \citep{gallerani:2018}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g., \citealt{Swinbank11a}) --- but the observed size scales could also be biased high due to the resolution limit.

>>>>

Summarizing, the virial parameter can be used to quantify the stability/boundedness of
%mm an MCC.
    a molecular cloud.
Accounting for the external pressure, a virial parameter of $\alpha_{\rm vir}\lesssim2$ would be unstable \citep{Bertoldi92b}.
%
Such a criterion is often used in \obs \citep[see e.g., ][]{Kauffmann17b}.

@@

In our stability analysis, we account for the combined effect of gas and stars \citep[derived exactly by][]{Rafikov01a}, and for the non-negligible disk thickness. This is done by adopting an approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modifies the $Q$ parameter for gas and stars by accounting for the vertical velocity dispersion
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the finite disk thickness reduces the gravity in the vertical direction, thereby making it easier for a system to maintain stability, and thus lowering the critical Toomre $Q_{\rm crit}$ from $\simeq$\,1 to 0.67 \citep{Goldreich65a}.
%
On the other hand, including the contribution of the stellar component promotes gravitational instability, and thus increases
$Q_{\rm crit}$, more-so if the stars have low velocity dispersion.
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

@@

Star formation is modeled using an H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar tracks from \ncode{starburst99} and include stellar feedback from supernovae (SNe) as well as winds from OB and asymptotic giant branch stars. Coupling to the gas is implemented via a sub-grid model for blastwaves, which accounts for radiative energy losses inside the cell. The remaining energy is injected into the ISM in kinetic and thermal energy forms. Radiation pressure on dust and gas is also included. Photoionization from local sources is neglected in this simulation \citep[see][for its effect]{Pallottini19a, Decataldo19a}.
%
Via feedback processes the gas can acquire energy both thermal and non-thermal form \citep{Agertz13a}; the latter is a term that mimics unresolved turbulence: such non-thermal energy is not affected by cooling but rather is dissipated on the eddy turn-over time scale \citep{Maclow99a}. See \citet{Pallottini17b} for further details.

The simulation zooms in on a galaxy named after the flower \flower whose properties are given in \citet{Pallottini17a} and are briefly summarized in the following. \flower is a Lyman-break galaxy that at $z\simeq 6$ is hosted by a dark matter halo of mass $M_{\rm DM}\simeq$\,10$^{11}$\,\Msun at the center of a cosmic web knot, and accretes mass from the IGM mainly via three filaments of length $\simeq$\,100\,kpc. \flower has a stellar mass of $M_\star\simeq$\,3\E{10}\,\Msun, a metallicity of $Z\simeq$\,0.5\,$Z_{\odot}$, a molecular gas mass of $M_{\rm H2}\simeq$\,5\E{7}\,\Msun, and a SFR of 30\,--\,80\,\Msun\,yr\pmOne.
%
The specific SFR of \flower is ${\rm sSFR} \equiv {\rm SFR}/M_{\star} \simeq$~4--40 ~Gyr\pmOne, compatible with the sSFR vs.\ $M_{\star}$ relation observed at high redshift \citep{Jiang16a}, which is higher than normal local galaxies. The effective stellar radius of \flower is $\sim$0.5\,kpc and the dark matter virial radius is $r_{\rm 200}\simeq$\,15\,kpc.

@@

The high pressure observed in \flower (see \citealt{Pallottini17a}) is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{Kohandel19a} and higher velocity accretion/SN-driven outflows \citep{Gallerani18a}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g., \citealt{Swinbank11a}) --- but the observed size scales could also be biased high due to the resolution limit.
